---
title: "Homework 3 Output - Using the Tidyverse"
author: Mike Maccia
format: pdf
editor: visual
---

Loading libraries

```{r}
library(tidyverse)
library(palmerpenguins)
```

## Task 1

### Question a: Why `read_csv` cannot be used to read the data.txt file

The function `read_csv()` can only be used to read in files that use the delimeters of commas or tabs. The function `read_csv2()` must be used in files with semicolons as the separator (commas can be used for decimal points).

```{r}
data <- read.csv2('~/ST558 Repo/Homework3/Data/data.txt',
                  header = T)
data
```

### Question b: Reading in 2nd file

In this file, "6" is the delimeter.

```{r}
data_2 <- read_delim('~/ST558 Repo/Homework3/Data/data2.txt',
                     delim = '6',
                     col_types= 'fdc')

data_2
```

## Task 2

Data tidying skills

### Question a: Reading Data

Reading in the `trailblazer.csv` data

```{r}
trailblazer <- read_csv('~/ST558 Repo/Homework3/Data/trailblazer.csv',
                   col_names= TRUE)
trailblazer
```

### Question b: Pivoting the data longer

```{r}
trailblazer_longer <- trailblazer |> 
    pivot_longer(cols = 2:11,
                 names_to = c('Game', 'Location'),
                 names_prefix = 'Game',
                 names_sep = '_',
                 values_to = 'Points') 

print(trailblazer_longer, n=5)

```

### Question c: Who scored more when playing at home versus away

```{r}

library(kableExtra)

trailblazer_wider <- trailblazer_longer |> 
    pivot_wider(names_from = (Location),
                 values_from = Points) |> 
    group_by(Player) |> 
    summarize(mean_home = mean(Home, na.rm=T),
              mean_away = mean (Away, na.rm=T)) |> 
    mutate(diff_points = (mean_home - mean_away)) |> 
    arrange(desc(diff_points)) |> 
    kbl(digits=2)

trailblazer_wider
```

While they did not necessarily score the most points, Jusuf Nurkic (6.67) and Robert Covington (6.5) scored on average more points at home than away through the first 10 games of the season.

## Task 3

### Question a. Describing what some values mean

<NULL> indicates that there were no values within a column. For example, there were no bill_length measurements for Gentoo species on Torgersen island.

\<dbl \[52\]\> indicates that within that cell there would are 52 observations (which are doubles) for bill length.

<list> indicates a list-column within a tribble. List-columns occur when each element within a column is a list. For example, the above cell of Adelie species on Torgersen island, there is a list of 52 doubles within that cell / element.

### Question b. Creating a new table

```{r}
penguins_island_ct <- penguins |> 
  select(species, island) |> 
  summarize(.groups = island)

penguins_island_ct
  
```

## Task 4

Replacing 2 missing values for bill length

```{r}
penguins_fixed_bill_length <- penguins |> 
  mutate(bill_length_mm = case_when(species == "Gentoo" & is.na(bill_length_mm) ~ 30, TRUE ~
                                      bill_length_mm )) |> 
  mutate(bill_length_mm = case_when(species == "Adelie" & is.na(bill_length_mm) ~ 26, TRUE ~
                                      bill_length_mm )) |> 
  arrange(bill_length_mm)

print(penguins_fixed_bill_length, n = 10)
```
